---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
```


# Reading in data


```{r}
vietnam <- read_csv("raw_data/thor_data_vietnam.csv",
                    col_types = cols(weapontypeclass = col_character(), tgtid = col_character(), 
                                     tgtcloudcover = col_character(), operationsupported = col_character(), 
                                     tgtcontrol = col_character(),tgtweather = col_character(), 
                                     releasefltspeed = col_character()))
```


```{r}
losses <- read_csv("data/us_aircraft_losses.csv")
```

```{r}
us_kills <- read_csv("data/us_kills.csv")
```

```{r}
losses_by_missile <- read_csv("data/losses_by_missile.csv")
```

```{r}
us_kills <- us_kills %>%
  mutate(date = ymd(date))
```



# Remove unneeded columns


```{r}
vietnam <- vietnam %>%
  select(-c(sourceid, sourcerecord, airforcegroup, airforcesqdn, releasealtitude, 
            releasefltspeed, resultsbda, tgtid, weapontypeclass, weaponsloadedweight,
            tgtweather))
```



# Data Wrangling and Visualisations

### total missions per day

```{r}
total_missions_daily <- vietnam %>%
  group_by(msndate) %>%
  summarise(num_missions = n()) %>%
  drop_na()
```

```{r}
total_missions_daily %>%
  ggplot() +
  aes(x = msndate, y = num_missions) %>%
  geom_line() 
```




### total missions per month

```{r}
total_missions_monthly <- vietnam %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month) %>%
  summarise(num_missions = n()) %>%
  drop_na()

total_missions_monthly <- total_missions_monthly %>%
  mutate(year_month = make_date(year, month)) 
```


```{r}
total_missions_monthly %>%
  ggplot() +
  aes(x = year_month, y = num_missions) %>%
  geom_line(col = "dark blue") +
  xlab("\nYear") +
  ylab("Number of Missions\n") +
  ggtitle("Total number of allied missions per month\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14))

```




### mission per day by service


```{r}
country_per_day <- vietnam %>%
  select(countryflyingmission, milservice, msndate) %>%
  group_by(msndate, countryflyingmission) %>%
  summarise(num_missions = n()) %>%
  drop_na()
  
```


```{r}
country_per_day %>%
  ggplot() +
  aes(x = msndate, y = num_missions, colour = countryflyingmission) +
  geom_line()
```




### missions per month by service


```{r}
country_per_month <- vietnam %>%
  select(countryflyingmission, milservice, msndate) %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month, countryflyingmission) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  mutate(year_month = make_date(year, month)) 
```


```{r}
country_per_month %>%
  ggplot() +
  aes(x = year_month, y = num_missions, colour = countryflyingmission) %>%
  geom_line() +
  xlab("\nYear") +
  ylab("Number of Missions\n") +
  ggtitle("Number of missions per month by all Allied Forces\n") +
  guides(colour = guide_legend(title = "Country Flying Mission")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(legend.title = element_text(size = 14))
```



### missions per day by US allies only

```{r}
country_per_day_allies <- vietnam %>%
  filter(countryflyingmission != "UNITED STATES OF AMERICA") %>%
  select(countryflyingmission, milservice, msndate) %>%
  group_by(msndate, countryflyingmission) %>%
  summarise(num_missions = n()) %>%
  drop_na()
```

```{r}
country_per_day_allies %>%
  ggplot() +
  aes(x = msndate, y = num_missions, colour = countryflyingmission) +
  geom_line()
```




### mission per month per US allies


```{r}
country_per_month_allies <- vietnam %>%
  filter(countryflyingmission != "UNITED STATES OF AMERICA") %>%
  select(countryflyingmission, milservice, msndate) %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month, countryflyingmission) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  mutate(year_month = make_date(year, month)) 
```

```{r}
country_per_month_allies %>%
  ggplot() +
  aes(x = year_month, y = num_missions, colour = countryflyingmission) %>%
  geom_line() +
  xlab("\nYear") +
  ylab("Number of Missions\n") +
  ggtitle("Number of missions per month by Non-US Forces\n") +
  guides(colour = guide_legend(title = "Country Flying Mission")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(legend.title = element_text(size = 14))
```



### mission per day per US service

```{r}
service_per_day_US <- vietnam %>%
  select(countryflyingmission, milservice, msndate) %>%
  filter(milservice == "USAF" | milservice == "USN" | milservice == "USMC" | milservice == "USMC") %>%
  group_by(msndate, milservice) %>%
  summarise(num_missions = n()) %>%
  drop_na()
```

```{r}
service_per_day_US %>%
  ggplot() +
  aes(x = msndate, y = num_missions, colour = milservice) +
  geom_line()
```



### mission per month per US service


```{r}
service_per_month_US <- vietnam %>%
  select(countryflyingmission, milservice, msndate) %>%
  filter(milservice == "USAF" | milservice == "USN" | milservice == "USMC" | milservice == "USMC") %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month, milservice) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  mutate(year_month = make_date(year, month)) 
```

```{r}
service_per_month_US %>%
  ggplot() +
  aes(x = year_month, y = num_missions, colour = milservice) %>%
  geom_line() +
  xlab("\nYear") +
  ylab("Number of Missions\n") +
  ggtitle("Number of missions per month by each US Military Service\n") +
  guides(colour = guide_legend(title = "US Military Service")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(legend.title = element_text(size = 14))
```



### mission per day by aircraft type

```{r}
aircraft_type_missions_daily <- vietnam %>%
  select(msndate, aircraft_root, milservice) %>%
  group_by(msndate, aircraft_root) %>%
  summarise(num_missions = n()) %>%
  drop_na()
  
```


### missions per month by aircraft type

```{r}
aircraft_type_missions_monthly <- vietnam %>%
  select(aircraft_root, milservice, msndate) %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month, aircraft_root) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  mutate(year_month = make_date(year, month)) 
```


### target country per day

```{r}
target_country_daily <- vietnam %>%
  select(msndate, tgtcountry) %>%
  filter(tgtcountry != "UNKNOWN" & tgtcountry != "WESTPAC WATERS") %>%
  group_by(msndate, tgtcountry) %>%
  summarise(num_missions = n()) %>%
  drop_na()
```

```{r}
target_country_daily %>%
  ggplot() +
  aes(x = msndate, y = num_missions, colour = tgtcountry) +
  geom_line()
```





### target country per month

```{r}
target_country_monthly <- vietnam %>%
  select(tgtcountry, msndate) %>%
  filter(tgtcountry != "UNKNOWN" & tgtcountry != "WESTPAC WATERS") %>%
  mutate(month = month(msndate)) %>%
  mutate(year = year(msndate)) %>%
  group_by(year, month, tgtcountry) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  mutate(year_month = make_date(year, month))
```

```{r}
target_country_monthly %>%
  filter(tgtcountry != "THAILAND", tgtcountry != "PHILLIPINES") %>% 
  ggplot() +
  aes(x = year_month, y = num_missions, colour = tgtcountry) %>%
  geom_line() +
  xlab("\nYear") +
  ylab("Number of Missions\n") +
  ggtitle("Number of missions per month by Target Country\n") +
  guides(colour = guide_legend(title = "Target Country")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(legend.title = element_text(size = 14))
```



### hull losses by service

```{r}
losses %>%
  filter(service == "USAF" | service == "USN" | service == "USMC" | service == "US Army") %>%
  group_by(type, service) %>%
  summarise(losses = sum(losses)) %>%
  ggplot() +
  aes(x = service, y = losses, fill = type) %>%
  geom_col(position = "dodge")
  
```



### hull losses by aircraft type

```{r}
losses %>%
  filter(service == "USAF" | service == "USN" | service == "USMC" | service == "US Army") %>%
  group_by(root, type) %>%
  arrange(desc(losses)) %>%
  filter(type == "Aircraft") %>%
  head(5) %>%
  ggplot() +
  aes(x = reorder(root, -(losses)), y = losses) +
  geom_col(fill = "dark blue")
```


### hull losses by helicopter type

```{r}
losses %>%
  filter(service == "USAF" | service == "USN" | service == "USMC" | service == "US Army") %>%
  group_by(root, type) %>%
  arrange(desc(losses)) %>%
  filter(type == "Helicopter") %>%
  head(5) %>%
  ggplot() +
  aes(x = reorder(root, -(losses)), y = losses) +
  geom_col(fill = "dark blue")
```




### US kills by aircraft type


```{r}
us_kills %>%
  group_by(aircraft) %>%
  summarise(num_kills = n()) %>%
  arrange(desc(num_kills)) %>%
  ggplot() +
  aes(reorder(x = aircraft, -(num_kills)), y = num_kills) +
  geom_col(fill = "dark blue")
```


### US kills by weapon

```{r}
us_kills %>%
  group_by(weapon) %>%
  summarise(num_kills = n()) %>%
  arrange(desc(num_kills)) %>%
  ggplot() +
  aes(reorder(x = weapon, -(desc(num_kills))), y = num_kills) %>%
  geom_col(fill = "dark blue") +
  coord_flip()
```


### US kills by Squadron

```{r}
us_kills %>%
  group_by(squadron) %>%
  summarise(num_kills = n()) %>%
  arrange(desc(num_kills)) %>%
  head(5) %>%
  ggplot() +
  aes(reorder(x = squadron, -desc(num_kills)), y = num_kills) +
  geom_col(fill = "dark blue") +
  coord_flip()
```



### US kills by enemy aircraft type


```{r}
us_kills %>%
  group_by(aircraft_destroyed) %>%
  summarise(num_kills = n()) %>%
  arrange(desc(num_kills)) %>%
  ggplot() +
  aes(reorder(x = aircraft_destroyed, -(num_kills)), y = num_kills) +
  geom_col(fill = "dark blue")
```


### US kills by month

```{r}
us_kills_monthly <- us_kills %>%
  mutate(month = month(date)) %>%
  mutate(year = year(date)) %>%
  group_by(year, month) %>%
  summarise(num_kills = n()) %>%
  mutate(year_month = make_date(year, month))
  
```

```{r}
us_kills_monthly %>%
  ggplot() +
  aes(x = year_month, y = num_kills) +
  geom_col()
```



### recoding target types

```{r}
target_type <- vietnam %>%
  select(milservice, msndate, valid_aircraft_root, tgttype, mfunc_desc, tgtcountry, periodofday, mfunc_desc_class) %>%
  mutate(tgttype = recode(tgttype, 
                          
                          #TRANSPORT  
                            "TRANSPORT\\TRK\\TRLR" = "TRUCKS", "TRUCK PARK" = "TRUCKS", "TRUCK PARK\\STOP" =
                            "TRUCKS", "TRANSPORT" = "TRUCKS", "TRUCK" = "TRUCKS", "TRUCK PRK ST" = "TRUCKS", 
                            "MOTOR VEHICLE" = "TRUCKS", "TRUCK PRK\\ST" = "TRUCKS", "VAN" = "TRUCKS",
                          #COMMUNICATIONS
                            "ANTENNA\\RADIO" = "COMMUNICATIONS", "RADIO COMMUNICATN" = "COMMUNICATIONS", 
                            "ANTENNA" = "COMMUNICATIONS", "RADIO" = "COMMUNICATIONS", "RADIO COMM ANY\\ANT" =
                            "COMMUNICATIONS", "COMMUNICATIONS\\ANY" = "COMMUNICATIONS", "COMMUNICATIONS VAN" =
                            "COMMUNICATIONS", "COMM AREA" = "COMMUNICATIONS",
                          #RADAR
                            "RADAR SRCH INTERCP" = "RADAR SITE", "RADAR VAN" = "RADAR SITE", "RADAR" = "RADAR SITE",
                          #HQ
                            "CCC\\HQ\\RADAR\\ETC" = "HEADQUARTERS", "COMMAND POST" = "HEADQUARTERS", "HQS AREA" =
                            "HEADQUARTERS", "HQ COMPLEX" = "HEADQUARTERS", "COMD CONTROL CENTR" = "HEADQUARTERS",
                            "BASE CAMP AREA" = "HEADQUARTERS", "MILITARY COMPLEX" = "HEADQUARTERS", "CAMP" = 
                            "HEADQUARTERS", "FIREBASE" = "HEADQUARTERS", "MIL BRKS\\ADMIN SPT" = "HEADQUARTERS",
                            "AREA COMPLEX" = "HEADQUARTERS", "BASE" = "HEADQUARTERS",
                          #GUN SITE
                            "85MM GUN SITE" = "GUN SITE", "130MM GUN SITE" = "GUN SITE", "155MM GUN SITE" = "GUN SITE",
                            "12.7 GUN SITE" = "GUN SITE", "57MM GUN SITE" = "GUN SITE", "105MM GUN SITE" = "GUN SITE",
                            "14.5MM GUN SITE" = "GUN SITE", "175MM GUN SITE" = "GUN SITE", "100MM GUN SITE" = 
                            "GUN SITE", "14.5MM GUN SITE" =  "GUN SITE", "NBR OF AAA PIECES" = "GUN SITE", 
                            "122MM GUN SITE" = "GUN SITE", "23MM GUN SITE" = "GUN SITE", "AAA\\37MM  CR MORE" = 
                            "GUN SITE", "GUN SITE" = "GUN SITE", "AAA SITE UNK CAL" = "GUN SITE", "ANTI-AIRCRAFT" =
                            "GUN SITE", "AA SITE" = "GUN SITE", "ARTILLERY" = "GUN SITE",
                          #SAM SITE
                            "SAMPANS" = "SAM SITE", "SAM CANISTER" = "SAM SITE", "SA\\3 SITE" = "SAM SITE",
                            "HH SAM SA\\7" = "SAM SITE", "MISSILE TRANSPORTER" = "SAM SITE", "SA SITE" = "SAM SITE",
                            "SAM LAUNCHER\\EQMT" = "SAM SITE", "MISSILE" = "SAM SITE",  "LAUNCHER (MISSILE)" = 
                            "SAM SITE", "SAM OXIDIZER" = "SAM SITE",
                          #RAILROAD
                            "MARINE RAILWAY" = "RAILROAD", "RAILROAD VEHICLES" = "RAILROAD", "RAILROAD COMPLEX" =
                            "RAILROAD", "RAILROAD SIDING" = "RAILROAD", "RAILROAD TRACK" = "RAILROAD",
                            "RAILROAD STA\\TERM" = "RAILROAD", "ROLLING STOCK" = "RAILROAD", "RAILROAD CROSSING" =
                            "RAILROAD", "RAILROAD YARDS" = "RAILROAD", "LOCOMT,RAIL VEH" = "RAILROAD", "ROLLING STOCK" =
                            "RAILROAD", "TRAIN CAR ANY" = "RAILROAD", "RAIL LINES\\ANY" = "RAILROAD",
                          #BRIDGE
                            "BRIDGE\\ANY" = "BRIDGE", "FOOT BRIDGE" = "BRIDGE", "RAILROAD BRIDGE" = "BRIDGE", 
                            "PONTOON BRIDGE" = "BRIDGE", "FORDS\\SUBSURFACE BRIDGE" = "BRIDGE", "BRIDGES" = "BRIDGE",
                          #AIRFIELD
                            "AIRCRAFT TRANSPORT" = "AIRFIELD", "AIRCRAFT ON GROUND" = "AIRFIELD", "AIRFIELD\\ANY TYPE" =
                            "AIRFIELD", "AIRFLD\\ACRFT\\ETC" = "AIRFIELD", "RUNWAY" = "AIRFIELD",
                          #PORT FACILITY
                            "SHIPYARD" = "PORT FACILITY", "WHARF" = "PORT FACILITY", "DRY DOCK" = "PORT FACILITY", 
                            "NAVAL BASE" = "PORT FACILITY", "NAVAL SUPPORT FACILITY" = "PORT FACILITY", 
                            "PIER" = "PORT FACILITY", "ANCHORAGE" = "PORT FACILITY",
                          #FACTORY
                            "FACTORY,ANY" = "FACTORY", "FACTORY INDUSTRIAL" = "FACTORY", "PLANT" = "FACTORY",
                          #POWER PLANT
                            "HYDROELEC PWR PLT" = "POWER PLANT", "ELEC. PWR. FAC" = "POWER PLANT", 
                            "THERMAL POWER PLANT" = "POWER PLANT", "COOLING TOWERS" = "POWER PLANT",
                            "ELEC. PWR. FAC." = "POWER PLANT",
                          #SUPPLY
                            "LOGISTICS\\ANY" = "SUPPLY", "OPEN STORAGE\\DUMP" = "SUPPLY", "SUPPLY STORAGE\\DMPS" =
                            "SUPPLY", "AMO\\POL\\SUP\\CRT" = "SUPPLY", "AREA\\DEPOT" = "SUPPLY",
                            "TRK\\PRK\\STORE AREA" = "SUPPLY", "FLOATING BAGS\\BOXES" = "SUPPLY",
                            "SUPPLIES (MISCELL.)" = "SUPPLY", "SUPPLIES\\RICE\\WEAPONS" = "SUPPLY", 
                            "AMMO STORAGE\\DUMP" = "SUPPLY", "BIVOUAC\\STORE AREA" = "SUPPLY", "CRATE\\STORAGE CONT" =
                            "SUPPLY", "STORAGE AREA" = "SUPPLY", "CIV\\DEF\\STOR" = "SUPPLY", "STORAGE AREA" =
                            "SUPPLY", "WEAPONS\\ANY" = "SUPPLY", "SILO" = "SUPPLY", "WAREHOUSE" = "SUPPLY",
                          #REFINERY
                            "REFINERY, ANY" = "REFINERY", "REFINERY, OIL\\GAS" = "REFINERY", "REFINERY (POL)" =
                            "REFINERY", "DRUM (POL)" = "POL",
                          #BOAT
                            "BOATS\\BARGES" = "BOAT", "WATER VEHICLES" = "BOAT", "WATERCRAFT" = "BOAT", "SHIP" =
                            "BOAT", "FERRY" = "BOAT", "GUNBOAT" = "BOAT",
                          #MINEFIELD
                            "MINEABLE AREA" = "MINEFIELD", "MINEFIELD (SEA)" = "MINEFIELD", "MINEFIELD (LAND)" =
                            "MINEFIELD",
                          #POL
                            "TANK (POL)" = "POL", "BUNKER FACILITY (POL)" = "POL", "POL BARRELS" = "POL",
                            "POL STORAGE" = "POL", "POL TANKS" = "POL",
                          #ROAD
                            "ROAD\\HIWAY\\APPRCH" = "ROAD/TRAIL", "ROAD CONST MATERIAL" = "ROAD/TRAIL",
                            "ROADS\\HIGHWAYS" = "ROAD", "TRAILS" = "ROAD/TRAIL", "ROAD" = "ROAD/TRAIL",
                          #TUNNEL
                            "RR TUNNEL" = "TUNNEL", "TUNNELS" = "TUNNEL",
                          #CAVE
                            "CAVES" = "CAVE", "CAVE STORAGE" = "CAVE",
                          #TROOPS
                            "PERSONNEL\\ANY" = "TROOPS", "FIGHTER" = "TROOPS", "SUSP ENEMY LOC" = "TROOPS", "SNIPER" =
                            "TROOPS", "BIVOUAC" = "TROOPS", "BIVUOAC AREA" = "TROOPS", "RPTD ENEMY LOC" = "TROOPS",
                            "FOX HOLES" = "TROOPS",
                          #DAM
                            "CAUSEWAY\\DAM" = "DAM", "CAUSEWAY" = "DAM", "DIKE" = "DAM",
                          #TOWER
                            "TOWER\\OUTPST\\LKOUT" = "TOWER",
                          #DEFENSIVE STRUCTURE
                            "TRENCHES" = "DEFENSE STRUCTURE", "BNKR\\TRENCH\\ETC" = "DEFENSIVE STRUCTURE",
                            "HIGHER\\REVETMENTS" = "DEFENSIVE STRUCTURE", "TRENCH" = "DEFENSIVE STRUCTURE",
                            "REVETMENT" = "DEFENSIVE STRUCTURE", "DEFENSE STRUCTURE" = "DEFENSIVE STRUCTURE",
                          #ARMORED VEHICLE
                            "TANK PARK" = "ARMORED VEHICLE", "TRACKED VEHICLE" = "ARMORED VEHICLE", "TANK"
                            = "ARMORED VEHICLE",
                          #CONSTRUCTION
                            "CONSTRUCTION VEHICLES" = "CONSTRUCTION", "CONSTRUCTION SITE" = "CONSTRUCTION",
                            "BULLDOZER" = "CONSTRUCTION",
                          #ROCKET SITE
                            "RCKT LNCHR (B40, ETC)" = "ROCKET SITE",
                          #CIV AREA
                            "VILLAGE" = "CIV AREA", "CIV POPULATN CENTR" = "CIV AREA",
                          #TRANSHIPMENT
                            "TRANSSHIP POINT" = "TRANSHIPMENT POINT",
                          #CANAL
                            "WATERWAY" = "CANAL",
                          #UNKNOWN
                            "UNKNOWN\\UNIDENTIFIED" = "UNIDENTIFIED"
                            ))
  
```



### enemy targets types

```{r}
target_type %>%
  group_by(tgttype) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(10) %>%
  ggplot() +
  coord_flip() +
  aes(reorder(x = tgttype, -desc(num_missions)), y = num_missions) +
  geom_col(fill = "dark blue") +
  xlab("Target Type\n") +
  ylab("\nNumber of Missions") +
  ggtitle("Top 10 Target Types\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text = element_text(size = 12))
  
```




### missions by type

```{r}
target_type %>%
  group_by(mfunc_desc) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(10) %>%
  ggplot() +
  aes(reorder(x = mfunc_desc, -(desc(num_missions))), y = num_missions) +
  coord_flip() +
  geom_col(fill = "dark blue") 
```


### targets by time of day

```{r}
target_type %>%
  mutate(periodofday = recode(periodofday, "D" = "Day", "N" = "Night", "M" = "Morning", "E" = "Evening")) %>%
  group_by(periodofday) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(4) %>%
  ggplot() +
  aes(reorder(x = periodofday, -(num_missions)), y = num_missions) +
  geom_col(fill = "dark blue")
  
  
```


###  Target types by aircraft root

```{r}
target_type %>%
  group_by(tgttype, valid_aircraft_root) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(9) %>%
  ggplot() +
  aes(reorder(x = tgttype, -(desc(num_missions))), y = num_missions, fill = valid_aircraft_root) +
  coord_flip() +
  geom_col(position = "dodge")
```


### types of mission

```{r}
target_type %>%
  group_by(mfunc_desc_class) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  ggplot() +
  aes(x = mfunc_desc_class, y = num_missions) +
  geom_col(fill = "dark blue")
  
```



### top 10 non-kinetic missions types

```{r}
target_type %>%
  filter(mfunc_desc_class == "NONKINETIC") %>%
  group_by(mfunc_desc) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(10) %>%
  ggplot() +
  aes(reorder(x = mfunc_desc, -(desc(num_missions))), y = num_missions) +
  coord_flip() +
  geom_col(fill = "dark blue") +
  xlab("Mission Type\n") +
  ylab("\nNumber of Missions") +
  ggtitle("Number of missions per non-kinetic mission type\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text = element_text(size = 12))
```




### top 10 kinetic misson types

```{r}
target_type %>%
  filter(mfunc_desc_class == "KINETIC") %>%
  group_by(mfunc_desc) %>%
  summarise(num_missions = n()) %>%
  drop_na() %>%
  arrange(desc(num_missions)) %>%
  head(10) %>%
  ggplot() +
  aes(reorder(x = mfunc_desc, -desc(num_missions)), y = num_missions) +
  coord_flip() +
  geom_col(fill = "dark blue") +
  xlab("Mission Type\n") +
  ylab("\nNumber of Missions") +
  ggtitle("Number of missions per kinetic mission type\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text = element_text(size = 12))
```


### Airbases

```{r}
vietnam %>%
  group_by(takeofflocation) %>%
  summarise(num_missions = n()) %>%
  arrange(desc(num_missions)) %>%
  head(10) %>%
  ggplot() +
  coord_flip() +
  aes(reorder(x = takeofflocation, -desc(num_missions)), y = num_missions) +
  geom_col(fill = "dark blue")
  
```





### ArcGIS master data

```{r}
arcgis_master <- vietnam %>% 
  select(msndate, countryflyingmission, milservice, aircraft_root, tgtcountry, 
         mfunc_desc_class, mfunc_desc, tgttype, callsign, 
         operationsupported, tgtlatdd_ddd_wgs84, tgtlonddd_ddd_wgs84) %>% 
  mutate(msndate = ymd(msndate)) %>%
  drop_na(msndate) %>%
  filter(tgtcountry != "UNKNOWN" & tgtcountry != "WESTPAC WATERS") 
```



### Radio relay missions


```{r}
laos <- arcgis_master %>%
  filter(tgtcountry == "LAOS") 
```


```{r}
#EC-121R Batcat

batcat <- laos[grep("BATCAT[0-9]*[a-zA-Z]*", laos$callsign, perl = TRUE), ]
num_batcat <- nrow(batcat)
```


```{r}
#EC-130 missions

num_ec130 <- nrow(laos %>% 
  filter(aircraft_root == "EC130") %>% 
  arrange(msndate))
```


```{r}
#QU-22 missions

num_qu22 <- nrow(laos %>% 
  filter(aircraft_root == "QU22") %>% 
  arrange(desc(msndate)))
```


```{r}
# total missions 

relay_missions <- data.frame(aircraft_type = c("EC-130", "QU-22", "EC-121R"),
                             number_missions = c(num_ec130, num_qu22, num_batcat))
relay_missions
```


```{r}
relay_missions %>%
  ggplot +
  aes(reorder(x = aircraft_type, -(number_missions)), y = number_missions) +
  geom_col(fill = "darkblue") +
  xlab("\nAircraft Type") +
  ylab("Number of Missions\n") +
  ggtitle("Number of Radio Relay and Command and Control missions in Laos\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text = element_text(size = 12))
  
```




```{r}
unique(arcgis_master$mfunc_desc)
```


### Psychological Warfare

```{r}
arcgis_master %>%
  filter(mfunc_desc == "PSYCHOLOGIC WARFARE") %>% 
  group_by(aircraft_root) %>%
  summarise(num_missions = n()) %>%
  arrange(desc(num_missions)) %>%
  head(5) %>%
  ggplot +
  aes(reorder(x = aircraft_root, -(num_missions)), y = num_missions) +
  geom_col(fill = "darkblue") +
  xlab("\nAircraft Type") +
  ylab("Number of Missions\n") +
  ggtitle("Number of psychological warfare missions by aircraft type\n") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) +
  theme(axis.title.x = element_text(size = 14)) +
  theme(axis.title.y = element_text(size = 14)) +
  theme(axis.text = element_text(size = 12))
```






############ ArcGIS data export ############



```{r}
# radio relay data

arcgis_relay <- arcgis_master %>% 
  filter(mfunc_desc == "RADIO RELAY") %>% 
  drop_na(tgtlonddd_ddd_wgs84)
```

```{r}
# extractions data

arcgis_extraction <- arcgis_master %>% 
  filter(mfunc_desc == "EXTRACTION (PLADS)" | mfunc_desc == "EXTRACTION (LAPES)" |
         mfunc_desc == "EXTRACTION (GPES)") %>% 
  drop_na(tgtlonddd_ddd_wgs84)
```

```{r}
# armed recce data

arcgis_recce <- arcgis_master %>%
  filter(mfunc_desc == "ARMED RECCE") %>%
  drop_na(tgtlonddd_ddd_wgs84)
```

```{r}
# kinetic missions 

arcgis_kinetic <- arcgis_master %>%
  filter(mfunc_desc_class == "KINETIC" | aircraft_root == "B52") %>% 
  drop_na(tgtlonddd_ddd_wgs84)
```

```{r}
#non-kinetic missions 

arcgis_nonkinetic <- arcgis_master %>%
  filter(mfunc_desc_class == "NONKINETIC" & aircraft_root != "B52") %>%
  drop_na(tgtlonddd_ddd_wgs84)
```

```{r}
#cambodia missions 

arcgis_cambodia <- arcgis_master %>%
  filter(tgtcountry == "CAMBODIA") %>%
  drop_na(tgtlatdd_ddd_wgs84)
```

```{r}
#laos missions 

arcgis_laos <- arcgis_master %>%
  filter(tgtcountry == "LAOS") %>%
  drop_na(tgtlatdd_ddd_wgs84) 
```


```{r}
#export to csv

write_csv(total_missions_monthly, "clean_data/monthly_missions.csv")
write_csv(arcgis_extraction, "clean_data/extraction.csv")
write_csv(arcgis_master, "clean_data/master.csv")
write_csv(arcgis_recce, "clean_data/recce.csv")
write_csv(arcgis_relay, "clean_data/relay_na.csv")
write_csv(arcgis_kinetic, "clean_data/kinetic.csv")
write_csv(arcgis_nonkinetic, "clean_data/nonkinetic.csv")
write_csv(arcgis_cambodia, "clean_data/cambodia.csv")
write_csv(arcgis_laos, "clean_data/laos.csv")
```



